Canonical rings of $\bq$-divisors on Minimal Rational Surfaces

### 2015-07-22 surfaces-write-up/surface-write-up.tex
edited intro and abstract

\begin{abstract}
 	Bounding generator and relation degrees of general $\bq$-divisors
	on projective spaces $\bp^n$ and Hirzebruch surfaces.
\end{abstract}

@@

\section{Introduction}
\label{sec:intro}
For any Weil $\bq$-divisor $D$ on a rational surface $X$, the graded
\textbf{canonical ring} $R(X, D) := \bigoplus_{d \geq 0} H^0(X, dD)$
has connections to rings of Hilbert and Siegel modular forms. In the
study of these canonical rings, it is natural to describe
presentations. We study the generators and relations of canonical
rings for projective spaces $\bp^m$ and Hirzebruch surfaces $\hirz_m$.
When $D$ is a general $\bq$-divisor on any $\bp^m$ or on any
$\hirz_m$, we give bounds on the generators and relations of
$R(X, D)$. In particular, we give a complete description of the
canonical ring when $X = \bp^m$ and $D$ is an $\bq$-divisor
containing one hypersurface and give stronger bounds when
$D$ is any effective $\bq$-divisor on $\bp^m$.

In part, this work is motivated by moving to generalize the work
of O'Dorney \cite{dorney:canonical} who gives similar descriptions
of canonical rings for general $\bq$-divisors on $\bp^1$. For
particular $\bq$ divisors on stacky curves, the work by
Voight--Zureick-Brown \cite{vzb:stacky} and Landesman--Ruhm--Zhang
\cite{lrz:spin-cring} on log spin canonical rings provides
useful bounds on rings of modular forms of Fuchsian groups.
The results of this work may have applications to Hilbert
and Siegel modular forms. \todo{reword last sentence}

\todo{Hassett-keel, see reference in davids paper about parametrizing $\alpha$ and considering stacky divisors. Would want to know whether they are a moduli space, though we just give generators and relations. This is one of the few examples other than that of nagata dealing with stacky surfaces.}

The work in this paper also has relations to certain rings of Hilbert 
modular forms and Siegel modular forms. Recall that Hilbert and Siegel 
modular forms are both to higher dimensions variables. In the case of 
Hilbert modular forms, one considers modular forms on a product of $m$
upper half planes, which transform under the action of finite index 
subgroups of the groups of positive determinant matrices. Similarly, Siegel 
modular forms are modular forms on the Siegel upper half space which are 
stable under the action of the symplectic group.

By a result by Voight and Zureick-Brown \cite[Proposition 6.1.5]{vzb:stacky}, 
one can translate the analytic category of 1-dimensional complex orbifolds to the algebraic category of stacky curves. 
We expect that  will similarly allow one to generalize yield an equivalence of
categories between an appropriate subcategory of $m$-dimensional orbifolds and
a subcategory of $m$-dimensional algebraic spaces.
In particular, this will allow us to translate rings of Hilbert modular forms
and Siegel modular forms whose underlying space is isomorphic to $\bp^m$ or a
Hirzebruch surface to canonical rings on the corresponding algebraic surfaces.
Thereby, our techniques would give a bound on generators and relations for
Hilbert and Siegel modular forms with certain underlying spaces.
\todo{include references}

Another closely related topic is the Hassett-Keel program 
initiated in \todo{was it actually initiated here? check vzb.} \cite{hassett:classical-and-minimal-models}, which aims to describe log canonical models of the form
\begin{align*}
	\overline {\mathscr M}_g(\alpha) := \bigoplus_{d \geq 0}H^0 \left( \overline {\mathscr M}_g, \lfloor d K_{\overline{\mathscr M}_g} + \alpha\delta \rfloor  \right) 
\end{align*}
in terms of certain moduli spaces. $\bq$-divisors on surfaces also naturally
appear when considering the canonical ring of surfaces of Kodaira dimension 1.
A surface has Kodaira dimension 1 if and only if it is an elliptic surface, and
their canonical ring are most naturally described in the setting of
$\bq$-divisors. \todo{Add references}

Furthermore, this work extends the study of canonical rings of
$\bq$-divisors on curves, as done by O'Dorney \cite{dorney:canonical},
Voight--Zureick-Brown \cite{vzb:stacky}, and Landesman--Ruhm--Zhang
\cite{lrz:spin-cring}, to general projective space and Hirzebruch
surfaces.

\todo{more!}

@@

\todo{put main results! use restate, maybe?}

\begin{thm}
\label{thm:proj-effective-intro}
Let $X = \bp^m$ with $\{x_0, \ldots, x_m\}$ a coordinate system for
$\bp^m$. Let $D = \sum_{i = 1}^{n} \alpha_i C_i$, with $\alpha_i =
\frac{p_i}{q_i} \in \bq$ and each $C_i$ a hypersurface of
$\bp^m$, be an effective divisor on $\bp^m$.

@@

\todo{Include summary of what each of the sections do.}

>>>>

\begin{abstract}
 	We give bounds on the degree of generation and relations of
	canonical rings of arbitrary $\bq$-divisors
	on projective spaces of all dimensions
	and Hirzebruch surfaces.
\end{abstract}

@@

\section{Introduction}
\label{sec:intro}
For any Weil $\bq$-divisor $D$ on a rational surface $X$, the graded
\textbf{canonical ring} is $R(X, D) := \bigoplus_{d \geq 0} H^0(X, \lfloor dD \rfloor)$.
We study the generators and relations of canonical
rings for projective spaces $\bp^m$ and Hirzebruch surfaces $\hirz_m$.
When $D$ is a general $\bq$-divisor on any $\bp^m$ or on any
$\hirz_m$, we give bounds on the generators and relations of
$R(X, D)$. In particular, we give a complete description of the
canonical ring when $X = \bp^m$ and $D$ is any effective
$\bq$-divisor.

In part, this work is motivated by
of O'Dorney \cite{dorney:canonical}, who gives similar descriptions
of canonical rings for general $\bq$-divisors on $\bp^1$. For
particular $\bq$ divisors on stacky curves, the work by
Voight--Zureick-Brown \cite{vzb:stacky} and Landesman--Ruhm--Zhang
\cite{lrz:spin-cring} on log spin canonical rings provides
useful bounds on rings of modular forms of Fuchsian groups.

The work in this paper may help calculate a presentation of certain rings of Hilbert 
modular forms and Siegel modular forms. Recall that Hilbert and Siegel 
modular forms are both generalizations of modular forms to higher dimensions.
By a result by Voight and Zureick-Brown \cite[Proposition 6.1.5]{vzb:stacky}, 
one can translate the analytic category of 1-dimensional complex orbifolds to
the algebraic category of stacky curves. 
We expect that one can generalize this result to yield an equivalence of
categories between an appropriate subcategory of $m$-dimensional orbifolds and
a subcategory of $m$-dimensional algebraic spaces.
In particular, this would allow us to translate rings of Hilbert modular forms
and Siegel modular forms whose coarse space is a projective space or a
Hirzebruch surface to canonical rings on the corresponding algebraic surfaces.
Thereby, our techniques would give a bound on generators and relations for
Hilbert and Siegel modular forms with certain underlying spaces.
\todo{include references}

Another closely related topic is the Hassett-Keel program 
initiated in \cite{hassett:classical-and-minimal-models}, which aims to describe log canonical models of the form
\begin{align*}
	\overline {\mathscr M}_g(\alpha) := \bigoplus_{d \geq 0}H^0 \left(
	\overline {\mathscr M}_g, \lfloor d K_{\overline{\mathscr M}_g} +
	\alpha\delta \rfloor  \right) 
\end{align*}
in terms of certain moduli spaces. Moreover, $\bq$-divisors on surfaces also naturally
appear when considering the canonical ring of surfaces of Kodaira dimension 1.
A surface has Kodaira dimension 1 if and only if it is an elliptic surface, and
their canonical ring are most naturally described in the setting of
$\bq$-divisors. \todo{Add references}

\todo{Aaron: This pararaph was already stated above. Can we remove it?}
This work extends the study of canonical rings of
$\bq$-divisors on curves, as done by O'Dorney \cite{dorney:canonical},
Voight--Zureick-Brown \cite{vzb:stacky}, and Landesman--Ruhm--Zhang
\cite{lrz:spin-cring}, to general projective space and Hirzebruch
surfaces.

\todo{more! Aaron: I think this is plenty of introduction}

@@

\todo{put main results! use restate, maybe?}
\todo{Aaron: Some of these theorem statements have been copied from older versions of the statements
and not updated to reflect the difference. I agree we should use Theorem-restate, but we should make
sure to copy the ones stated below, and not copy these ones}
\begin{thm}
\label{thm:proj-effective-intro}
Let $X = \bp^m$ with $\{x_0, \ldots, x_m\}$ a coordinate system for
$\bp^m$. Let $D = \sum_{i = 1}^{n} \alpha_i C_i$, with $\alpha_i =
\frac{p_i}{q_i} \in \bq$ and each $C_i$ a hypersurface of
$\bp^m$, be an effective divisor on $\bp^m$.

@@

\todo{Include summary of what each of the sections do.}0
